An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-11 VLDB2026/main.tex
Add files via upload
Add caps Experiment 2-2

	\setlength{\textfloatsep}{0cm}
	\setlength{\floatsep}{0cm}
	\begin{table*}[t]
		\centering
		%\setlength{\belowcaptionskip}{-0.3cm}
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\caption{Comparison of AF-ANN search algorithms}
		\small
		\label{tab:compair_1}
		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
			\hline
			NHQ & Graph & C & Y & N& N& N & N & N \\
			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
			CAPS & IVF & C & Y & N & N & N & Y & Y \\
			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\

@@

	\noindent\textbf{\underline{UNG \cite{UNG}.}} Graph-based methods (NHQ, Filtered-DiskANN, ACO-\\RN) do not guarantee result completeness, while CAPS lacks robustness when base and query vectors differ in attribute cardinality. UNG is proposed to overcome both limitations.

@@

	In database systems (VBASE, PASE, Milvus), when the system builds both B+ tree and vector indexes, queries typically utilize only the B+ tree. Faiss offers an ID filtering mechanism, it operates solely during query execution and does not influence index building. CAPS requires that the number of attributes in a query match those used during index construction. ACORN-1 and ACORN-\(\gamma\) are specifically designed for single-attribute indexing. Hence, they are excluded from this comparison.

@@

Furthermore, current attribute-filtering algorithms offer limited support for complex filtering conditions. For example, Filtered-DiskANN  supports single-attribute filtering, and in multi-attribute scenarios, it only supports OR conditions between attributes. 
%The original implementations of 
NHQ and CAPS impose strict constraints on the number of base and query attributes and only support AND conditions. While UNG is relatively flexible—supporting both AND and OR logic—it still lacks support for arbitrary Boolean expressions (e.g., combinations of AND and OR). The lack of flexible Boolean filtering remains a key challenge, limiting the practicality of hybrid query methods in real-world systems with diverse and complex query requirements.

>>>>

	\setlength{\textfloatsep}{0cm}
	\setlength{\floatsep}{0cm}
	\begin{table*}[t]
		\centering
		%\setlength{\belowcaptionskip}{-0.3cm}
		\setlength{\abovecaptionskip}{0.1cm}
		\setlength{\belowcaptionskip}{-0.2cm}
		\caption{Comparison of AF-ANN search algorithms}
		\small
		\label{tab:compair_1}
		\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
			\hline
			\textbf{Algorithm} & \textbf{Base} & \textbf{Filter Type} & \textbf{AND} & \textbf{OR} & \textbf{Flexible Attributes} & \textbf{Complex Boolean} & \textbf{Dynamic Insert} & \textbf{Multi Thread} \\
			\hline
			NHQ & Graph & C & Y & N& N& N & N & N \\
			FilteredVamana & Graph & C & N & Y & Y & N & Y & Y \\
			StitchedVamana & Graph & C & N & Y & Y & N & N & Y  \\
			CAPS & IVF & C & Y & N & Y & N & Y & Y \\
			ACORN & Graph & B & Y & Y & Y & Y & Y & Y \\
			UNG & Graph & B & Y & Y & Y & N & Y & Y \\ 
			Puck & IVF & B & Y & Y & Y & N & Y & Y \\
			% ParlayANN & Graph+IVF & B & N & N & Y & N & N & Y \\

@@

	\noindent\textbf{\underline{UNG \cite{UNG}.}} Graph-based methods (NHQ, Filtered-DiskANN, ACO-\\RN) do not guarantee result completeness. UNG is proposed to overcome both limitations.

@@

	In database systems (VBASE, PASE, Milvus), when the system builds both B+ tree and vector indexes, queries typically utilize only the B+ tree. Faiss offers an ID filtering mechanism, it operates solely during query execution and does not influence index building. ACORN-1 and ACORN-\(\gamma\) are specifically designed for single-attribute indexing. Hence, they are excluded from this comparison.

@@

Furthermore, current attribute-filtering algorithms offer limited support for complex filtering conditions. For example, Filtered-DiskANN  supports single-attribute filtering, and in multi-attribute scenarios, it only supports OR conditions between attributes. 
%The original implementations of 
NHQ imposes strict constraints on the number of base and query attributes and only support AND conditions. While UNG is relatively flexible—supporting both AND and OR logic—it still lacks support for arbitrary Boolean expressions (e.g., combinations of AND and OR). The lack of flexible Boolean filtering remains a key challenge, limiting the practicality of hybrid query methods in real-world systems with diverse and complex query requirements.
